---
title: "TE_computation"
date: "Last update: Dec 2nd 2020"
output: html_document
---

# Preparations
## load libraries
```{r setup, include=FALSE}

library(here)
library(tidyverse)
library(tidylog)
library(ggpubr)
#library(regclass) # for correlation
#library(effects)
# library(jtools)
#library(FRACTION) # this allows us to use the is.wholenumber function

`%notin%` <- Negate(`%in%`)

```


# Read in data
```{r}
TE_computation <- read.csv(here::here("computational model/TE_computation.csv"))
```


# Simulation 1
## Infants of the same developmental level with different word vocabularies and balances of exposure
### V = 600
```{r}
Simulation1 <- TE_computation %>%
  # V = 600
  filter(V == 600) %>%
  # Convert to long data
  pivot_longer(c(W, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type == c("W", "D", "N")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
  mutate(vocab_type = recode(vocab_type, 
                             W = "Panel A: Total vocabulary (W)", 
                             D = "Panel B: Dominant vocabulary (D)", 
                             N = "Panel C: Non-dominant vocabulary (N)")) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  #mutate(bin_W = cut_interval(W, 12,
  #                   labels = c(100,200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200))) %>%
  ggplot(aes(x = number, y = TE, color = bin_B)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  ylim(0, 600) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Number of words", 
       y = "Total number of translation equivalents (TE)", 
       color = "Balance (B)"
       #, title = "Number of translation equivalent by balance in relation to different input measures "
       ) +
  guides(colour = guide_legend(nrow = 1))

ggsave("plot_Simulation1.png", plot_inputs_TE,
       width = 10, height = 5)

```


# Simulation 2
## Patterns of translation equivalent learning as a function of different developmental timepoints
```{r}
timepoints <- c(150, 300, 450, 600)

Simulation2 <- TE_computation %>%
  filter(V %in% timepoints) %>%
  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %notin% c("D", "N", "C")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("Nu", "Du", "TE"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  group_by(V, bin_B, vocab_type) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ V) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
  #scale_y_continuous(limits = c(0, 1200)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom",
        plot.title = element_text(hjust = 0.5)) + 
  labs(x = "Balance", 
       y = "Number of concepts", 
       title = "Vocabulary composition across different developmental timepoints (V)") 

ggsave("plot_Simulation2.png", plot_composition_byV,
        width = 10, height = 5)

```


# Simulation 3
## Non-independence of dominant-language (D) and non-dominant language words (N)
```{r}

# I * (D*N/V)
I1 = 1.5
I2 = 0.5

Simulation3 <- TE_computation %>%
  filter(V==600) %>%
  ## Calculate where I > 1
  mutate(TE_easier = I1*(D*N/V)) %>%
  ## Calculate where I < 1
  mutate(TE_harder = I2*(D*N/V)) %>%
  pivot_longer(c(TE, TE_harder, TE_easier), 
               names_to = "Model", values_to = "number_TE") %>%
  mutate(Model = factor(Model, levels=c("TE_harder", "TE", "TE_easier"))) %>%
  mutate(Model = recode(Model, 
                        TE = "I = 1 (assuming independence of D and N)",
                        TE_easier = "I = 1.5 (TEs are more easily learned)",
                        TE_harder = "I = 0.5 (TE are less easily learned)")) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = W, y = number_TE, color = bin_B)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  theme_minimal() + 
  facet_grid(. ~ Model) +
  ylim(0, 600) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Total word vocabulary (W)", 
       y = "Total number of translation equivalents (TE)", 
       color = "Balance (B)") +
  guides(colour = guide_legend(nrow = 1))

ggsave("plot_Simulation3.png", Simulation3, 
       width = 10, height = 5)



#Simulation3 <- TE_computation %>%
#  pivot_longer(c(TE_harder, TE_easier), 
#               names_to = "Nonindependence_type", values_to = "Simulated_TEs") %>%
#  mutate(Nonindependence_type = recode(Nonindependence_type, 
#                                       TE_easier = "I = 1.5 (TEs are more easily learned)",
#                                       TE_harder = "I = 0.5 (TE are less easily learned)")) %>%
#  ggplot(aes(y = Simulated_TEs, x = TE, color = Nonindependence_type)) +
#  geom_smooth(method = "lm", se = F) +
#  geom_point(shape = 23) +
#  geom_abline(linetype = 2, alpha = 0.5) +
#  labs(y = "Simulated number of TEs assuming non-independence of D and N", 
#       x = "Expected number of TEs assuming independence of D and N",
#       color = "Scenarios"
#       #, title = "Non-independence vs. Independence of dominant and non-dominant words"
#       ) +
#  theme(plot.title = element_text(hjust = 0.5)) +
#  theme_minimal()

##ggsave("plot_Simulation3.png", Simulation3, 
##       width = 8, height = 6)

```


















# Plotting relations between variables
## Stacked barchart showing the vocabulary composition across different balance proportions
```{r}

#plot_composition <- TE_computation %>%
#  # Convert to long data
#  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
#  filter(vocab_type %notin% c("D", "N", "C")) %>%
#  mutate(vocab_type = factor(vocab_type, levels=c("TE", "Du", "Nu"))) %>%
#  mutate(bin_W = cut(W, breaks = 3, 
#                     labels = c("0-400", "401-800", "801-1200"))) %>%
#  group_by(bin_W, vocab_type, B) %>%
#  summarise(n_words = mean(number)) %>%
#  ggplot(aes(x=B, y = n_words, fill = vocab_type)) + 
#  geom_bar(position="stack", stat="identity") +
#  facet_grid(. ~ bin_W) +
#  #scale_y_continuous(limits = c(0, 1200)) +
#  theme_minimal() +
#  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") + 
#  labs(x = "Balance", 
#       y = "Number of words") 

plot_composition_byB <- TE_computation %>%
  # Convert to long data
  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %notin% c("D", "N", "C")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("Nu", "Du", "TE"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  #mutate(bin_W = cut_interval(W, 12,
  #                   labels = c(100,200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200))) %>%
  mutate(bin_W = cut_interval(W, 8,
                     labels = c(150, 300, 450, 600, 750, 900, 1050, 1200))) %>%
  group_by(bin_B, vocab_type, bin_W) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x = bin_W, y = n_words, fill = vocab_type)) + 
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ bin_B) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
  #scale_y_continuous(limits = c(0, 1200)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom",
        plot.title = element_text(hjust = 0.5)) +
  labs(x = "Total vocabulary size", 
       y = "Number of concepts", 
       title = "Vocabulary composition across different balance proportions") 

ggsave("plot_composition_byB.png", plot_composition_byB)




plot_composition_byW <- TE_computation %>%
  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %notin% c("D", "N", "C")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("Nu", "Du", "TE"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  mutate(bin_W = cut_interval(W, 6,
                     labels = c(200, 400, 600, 800, 1000, 1200))) %>%
  group_by(bin_B, vocab_type, bin_W) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ bin_W) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
  #scale_y_continuous(limits = c(0, 1200)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") + 
  labs(x = "Balance", 
       y = "Number of concepts", 
       title = "Vocabulary composition across different vocabulary size") 

plot_composition_byV <- TE_computation %>%
  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %notin% c("D", "N", "C")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("Nu", "Du", "TE"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  group_by(V, bin_B, vocab_type) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ V) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
  #scale_y_continuous(limits = c(0, 1200)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom",
        plot.title = element_text(hjust = 0.5)) + 
  labs(x = "Balance", 
       y = "Number of concepts", 
       title = "Vocabulary composition across different developmental timepoints (V)") 

ggsave("plot_composition_byV.png", plot_composition_byV,
        width = 10, height = 5)

```

## Compare different vocabulary measures across balance
```{r}
vocab_label <- c("Concept Vocabulary (C)", "Unique Dominant Words (Du)", "Unique Non-dominant Words (Nu)")

plot_words <- TE_computation %>%
  # Convert to long data
  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %notin% c("D", "N")) %>%
  mutate(vocab_type = as.factor(vocab_type)) %>%
  mutate(vocab_type = recode(vocab_type, C = "Concept vocabulary (C)", Du = "Unique dominant words (Du)",
         Nu = "Unique non-dominant words (Nu)", TE = "Translation equivalents")) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  #mutate(bin_W = cut_interval(W, 12,
  #                   labels = c(100,200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200))) %>%
  group_by(bin_B, vocab_type, W) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x = W, y = n_words, color = bin_B)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  ylim(0, 600) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Total vocabulary Size (W)", 
       y = "Number of words", 
       color = "Balance",
       title = "Word vocabulary size by balance under different vocabulary measures") +
  guides(colour = guide_legend(nrow = 1))

  
ggsave("plot_words.png", plot_words,
       width = 10, height = 5)



plot_inputs <- TE_computation %>%
  # Convert to long data
  pivot_longer(c(W, D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type == c("W", "D", "N", "TE")) %>%
  mutate(vocab_type = as.factor(vocab_type)) %>%
  mutate(vocab_type = fct_relevel(vocab_type, "W")) %>%
  mutate(vocab_type = recode(vocab_type, W = "Total vocabulary (W)", D = "Dominant vocabulary (D)", 
                             N = "Non-dominant vocabulary (N)", TE = "Translation equivalents (TE)")) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  #mutate(bin_W = cut_interval(W, 12,
  #                   labels = c(100,200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200))) %>%
  group_by(V, bin_B, vocab_type) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x = V, y = n_words, color = bin_B)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  #ylim(0, 600) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Different timepoints = Learnable vocaobulary (V)", 
       y = "Number of words", 
       color = "Balance",
       title = "Word vocabulary size by balance under different input measures and translation equivalents") +
  guides(colour = guide_legend(nrow = 1))

ggsave("plot_inputs.png", plot_inputs,
       width = 10, height = 5)







plot_inputs_TE_V <- TE_computation %>%
  filter(W == 250) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = V, y = TE, color = bin_B)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  theme_minimal() + 
  #facet_grid(. ~ bin_W) +
  #ylim(0, 600) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Developmental timepoint (V)", 
       y = "Total number of translation equivalents (TE)", 
       color = "Balance (B)") +
  guides(colour = guide_legend(nrow = 1))


plot_inputs_TE_V <- TE_computation %>%
  filter(W == 400) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  #mutate(bin_W = cut_interval(W, 3,
  #                   labels = c("W = 0 - 400" , "W = 401 - 800", "W = 801 - 1200"))) %>%
  ggplot(aes(x = V, y = TE, color = bin_B)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  theme_minimal() + 
  #facet_grid(. ~ bin_B) +
  #ylim(0, 600) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Developmental timepoint (V)", 
       y = "Total number of translation equivalents (TE)", 
       color = "Balance (B)") +
  guides(colour = guide_legend(nrow = 1))

ggsave("plot_inputs_TE_V.png", plot_inputs_TE_V,
       width = 6, height = 6)


plot_inputs_TE_V <- TE_computation %>%
  #filter(W == 100) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  mutate(bin_W = cut_interval(W, 4,
                     labels = c("W = 0 - 300" , "W = 301 - 600", "W = 601 - 900", "W = 901 - 1200"))) %>%
  ggplot(aes(x = V, y = TE, color = bin_W)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  theme_minimal() + 
  facet_grid(. ~ bin_B) +
  #ylim(0, 600) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Developmental timepoint (V)", 
       y = "Total number of translation equivalents (TE)", 
       color = "Balance (B)") +
  guides(colour = guide_legend(nrow = 1))

ggsave("plot_inputs_TE_V.png", plot_inputs_TE_V,
       width = 10, height = 5)

```

## Plotting different scenarios
```{r}
plot_scenarios <- TE_computation %>%
  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type == c("D", "N", "TE")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("N", "D", "TE"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  group_by(V, bin_B, vocab_type) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ V) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
  #scale_y_continuous(limits = c(0, 1200)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") + 
  labs(x = "Balance", 
       y = "Number of words (W)", 
       title = "Vocabulary composition across different developmental timepoints (V)") 

```

## Plotting relations between TEs and different variables
##### Relation between TE & W, D, & N #####
```{r}
### Plot: Y = TE, X = D / N
plot_TEandWDN <- TE_computation %>%
  pivot_longer(c(W, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("W", "D", "N")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
  ggplot(aes(x = number, y = TE)) +
  facet_grid(. ~ vocab_type) +
  labs(x = "Words produced", y = "Total number of TEs") +
  #geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_y_continuous(limits = c(0, 600)) +
  theme_minimal()

ggsave("plot_TEandWDN.png", plot_TEandWDN)


### Plot: Y = TE, X = D / N, color by B
plot_TEandWDN_byB <- TE_computation %>%
  pivot_longer(c(W, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("W", "D", "N")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = number, y = TE, colour = bin_B)) +
  facet_grid(. ~vocab_type) +
  labs(x = "Words produced", y = "Total number of TEs",
       colour = "Bins of balance") +
  #geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_x_continuous(limits = c(0, 1200)) +
  scale_y_continuous(limits = c(0, 600)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1))

ggsave("plot_TEandWDN_byB.png", plot_TEandWDN_byB)
```

##### Relation between W & TE, Du, & Nu #####
```{r}
### Plot: Y = W, X = TE/ Du / Nu
plot_WandTEDuNu <- TE_computation %>%
  pivot_longer(c(TE, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("TE", "Du", "Nu")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("TE", "Du", "Nu"))) %>%
  #mutate(bin_B = cut(B, breaks = 6, 
  #                   labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = number, y = W)) +
  facet_grid(. ~vocab_type) +
  labs(x = "Words produced", y = "Total vocabulary size") +
  #geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_y_continuous(limits = c(0, 1200)) +
  theme_minimal()

ggsave("plot_WandTEDuNu.png", plot_WandTEDuNu)


### Plot: Y = W, X = TE/ Du / Nu, color by B
plot_WandTEDuNu_byB <- TE_computation %>%
  pivot_longer(c(TE, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("TE", "Du", "Nu")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("TE", "Du", "Nu"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = number, y = W, colour = bin_B)) +
  facet_grid(. ~vocab_type) +
  labs(x = "Words produced", y = "Total vocabulary size",
       colour = "Bins of balance") +
  #geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_y_continuous(limits = c(0, 1200)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1))

ggsave("plot_WandTEDuNu_byB.png", plot_WandTEDuNu_byB)
```

##### Relation between C & W, D, & N #####
```{r}
### Plot: Y = C, X = W / D / N
plot_CandWDN <- TE_computation %>%
  pivot_longer(c(TE, W, D, N, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("W", "D", "N")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
  #mutate(bin_B = cut(B, breaks = 6, 
  #                   labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = number, y = C)) +
  facet_grid(. ~vocab_type) +
  labs(x = "Words produced", y = "Number of concepts") +
  #geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_y_continuous(limits = c(0, 600)) +
  theme_minimal()

ggsave("plot_CandWDN.png", plot_CandWDN)


### Plot: Y = C, X = W / D / N, color by B
plot_CandWDN_byB <- TE_computation %>%
  pivot_longer(c(TE, W, D, N, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("W", "D", "N")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = number, y = C, colour = bin_B)) +
  facet_grid(. ~vocab_type) +
  labs(x = "Words produced", y = "Number of concepts",
       colour = "Bins of balance") +
  #geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_x_continuous(limits = c(0, 1200)) +
  scale_y_continuous(limits = c(0, 600)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1))

ggsave("plot_CandWDN_byB.png", plot_CandWDN_byB)
```


# Plotting relations between variables (while holding some variables constant)
## TE ~ predictors (input)
```{r}
#TE_Constant <- lm(TE ~ ., data = TE_computation) 
#summary(TE_Constant)

#### run a regression with TE ~ B, D, N
TE_InputConstant <- lm(TE ~ 0 + B + D + N, data = TE_computation) 
summary(TE_InputConstant)

### plot the relation between TE & B / D / N
#effect_plot(TE_InputConstant, pred = B, plot.points = TRUE)
effect_B <- effect(term = "B", mod = TE_InputConstant) %>%
  data.frame()

plotConstant_TE_B <- ggplot(effect_B, aes(x = B, y = fit)) +
  geom_line() +
  ylim(-25, 600) +
  labs(x = 'Balance', y = 'Number of TEs') +
  theme_minimal()

#ggsave("plotConstant_TE_B.png", plotConstant_TE_B)

### plot the relation between TE & D
#effect_plot(TE_InputConstant, pred = D, plot.points = TRUE)
effect_D <- effect(term = "D", mod = TE_InputConstant) %>%
  data.frame()

plotConstant_TE_D <- ggplot(effect_D, aes(x = D, y = fit)) +
  geom_line() +
  ylim(-25, 600) +
  labs(x = 'Dominant vocabulary', y = 'Number of TEs') +
  theme_minimal() 

#ggsave("plotConstant_TE_D.png", plotConstant_TE_D)

### plot the relation between TE & N
#effect_plot(TE_InputConstant, pred = N, plot.points = TRUE)
effect_N <- effect(term = "N", mod = TE_InputConstant) %>%
  data.frame()

plotConstant_TE_N <- ggplot(effect_N, aes(x = N, y = fit)) +
  geom_line() +
  ylim(-25, 600) +
  labs(x = 'Non-dominant vocabulary', y = 'Number of TEs') +
  theme_minimal()

#ggsave("plotConstant_TE_N.png", plotConstant_TE_D)
  
## put the three plots together
plotConstant_TE_BDN <- ggarrange(plotConstant_TE_B, plotConstant_TE_D, plotConstant_TE_N, 
                                 ncol = 3, nrow = 1,
                                 widths = 3, heights = 1) %>%
  annotate_figure(top = text_grob("linear regression model: TE ~ 0 + B + D + N", face = "bold", size = 14))

ggsave("plotConstant_TE_BDN.png", plotConstant_TE_BDN, 
       width = 10, height = 5)
```


## TE ~ predictors (output)
```{r}
#### run a regression with TE ~ C, Du, Nu
TE_OutputConstant <- lm(TE ~ 0 + C + Du + Nu, data = TE_computation) 
summary(TE_OutputConstant)

## plot the relation between TE & C
#effect_plot(TE_OutputConstant, pred = C, plot.points = TRUE)
effect_C <- effect(term = "C", mod = TE_OutputConstant) %>%
  data.frame()

plotConstant_TE_C <- ggplot(effect_C, aes(x = C, y = fit)) +
  geom_line() +
  labs(x = 'Concept vocabulary', y = 'Number of TEs') +
  ylim(-300, 600) +
  theme_minimal()

#ggsave("plotConstant_TE_C.png", plotConstant_TE_C)

## plot the relation between TE & Du
#effect_plot(TE_OutputConstant, pred = Du, plot.points = TRUE)
effect_Du <- effect(term = "Du", mod = TE_OutputConstant) %>%
  data.frame()

plotConstant_TE_Du <- ggplot(effect_Du, aes(x = Du, y = fit)) +
  geom_line() +
  labs(x = 'Unique dominant vocabulary', y = 'Number of TEs') +
  ylim(-300, 600) +
  theme_minimal()

#ggsave("plotConstant_TE_Du.png", plotConstant_TE_Du)

## plot the relation between TE & Nu
#effect_plot(TE_OutputConstant, pred = Nu, plot.points = TRUE)
effect_Nu <- effect(term = "Nu", mod = TE_OutputConstant) %>%
  data.frame()

plotConstant_TE_Nu <- ggplot(effect_Nu, aes(x = Nu, y = fit)) +
  geom_line() +
  labs(x = 'Unique non-dominant vocabulary', y = 'Number of TEs') +
  ylim(-300, 600) +
  xlim(0,600) +
  theme_minimal()

#ggsave("plotConstant_TE_Du.png", plotConstant_TE_Du)

## put the three plots together
plotConstant_TE_CDuNu <- ggarrange(plotConstant_TE_C, plotConstant_TE_Du, plotConstant_TE_Nu, 
                                 ncol = 3, nrow = 1,
                                 widths = 3, heights = 1) %>%
  annotate_figure(top = text_grob("linear regression model: TE ~ 0 + C + Du + Nu", face = "bold", size = 14))

ggsave("plotConstant_TE_CDuNu.png", plotConstant_TE_CDuNu, 
       width = 10, height = 5)
```

# Estimating  if the models is biased towards ot against learning TEs
```{r}
## plot
plot_learnVSpredict <- TE_computation %>%
  mutate(predicted_TE = TE, #predicted TE is based on D*N/V (where V is the max. possible words)
         learned_TE = (D*N)/(W/2)) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = learned_TE, y = predicted_TE, color = bin_B)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  labs(x = "Learned translation equivalents (= (D*N)/(W/2))",
       y = "Predicted translation equivalents under the model (= (D*N)/V)",
       color = "Bins of balance") +                                                    
  theme_minimal()

ggsave("plot_learnVSpredict.png", plot_learnVSpredict)

## Kolmogorov-Smirnov test to compare the observed and predicted TE
TE_computation <- TE_computation %>%
  mutate(predicted_TE = TE, #predicted TE is based on D*N/V (where V is the max. possible words)
         learned_TE = (D*N)/(W/2))

stats::ks.test(TE_computation$predicted_TE, TE_computation$learned_TE)

```

